The error in Bayesian phylogenetic reconstruction when speciation co-occurs

### 2019-11-25 razzo-methods.tex
Describe how we picked the parameter settings

In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
to the common per-species speciation and extinction rates 
already present in the standard BD model. 
Additionally, MBD relies on two additional parameters, $\nu$ and $q$. 
The first, $\nu$, is the rate at which an environmental change is triggered.
When such event is triggered, each species present in the phylogeny 
at that moment has a probability $q$ to speciate at that time.
This kind of speciation is of a different nature with respect to 
the one triggered by $\lambda$. 
In fact, whereas parameter $\lambda$ can be seen as 
describing a sympatric process, $\nu$ induces the rise of 
geographical barriers that interrupt the gene flow,
resulting in allopatric speciation. 
Even though multiple speciations can co-occur, 
polytomies are not allowed in such process as each species can speciate only 
once during a co-occuring speciation event.
A likelihood expression for the process is provided in \cite{mbd}.

>>>>

In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
to the common per-species speciation and extinction rates 
already present in the standard BD model. 
Additionally, MBD relies on two additional parameters, $\nu$ and $q$. 
The first, $\nu$, is the rate at which an environmental change is triggered.

When such event is triggered, each species present in the phylogeny 
at that moment has a probability $q$ to speciate at that time.
This kind of speciation is of a different nature with respect to 
the one triggered by $\lambda$. 
In fact, whereas parameter $\lambda$ can be seen as 
describing a sympatric process, $\nu$ induces the rise of 
geographical barriers that interrupt the gene flow,
resulting in allopatric speciation. 
Even though multiple speciations can co-occur, 
polytomies are not allowed in such process as each species can speciate only 
once during a co-occuring speciation event.

A likelihood expression for the process is provided in \cite{mbd}.

\subsection{Parameter settings}

We ran multiple pilot experiments with 1 replicate to arrive at our final
parameter settings. We devised a set of rules to make a verdict about the
settings \richel{(script at \url{https://github.com/richelbilderbeek/razzo_project/blob/master/scripts/90_collect_run_times.R},
resulting verdicts at \url{https://github.com/richelbilderbeek/razzo_project/blob/master/verdict.md})}:
\itemize{
  \item quality: 95\% of all individual runs should have an ESS of at least 200.
  \item feasibility: 95\% of all individual runs should finish within 10 days.
  \item reproducibility: the mean run-time of all finished runs should be less than 24 hours 
    \richel{I suggest}.
  \item relevance: the percentage of taxa created by the MB process should be
    as high as possible
}

We searched through parameter space until these criteria were met.
All the parameter settings used in the pilot experiments can be found at 
\url{https://github.com/richelbilderbeek/razzo_project/blob/master/overview.md}.
Due to the low number of replicates, we were unable (nor tried)
to draw conclusions based on the results.
